Adapting the MVDR to a GPU for active sonar systems of different sizes.

### 2013-01-23 ASA2013/asa2013.tex
la2pdf commit - 2013:01:23	02:05:28	(Wed23.Jan)

Although LCA promotes good image quality, its two most attractive traits is perhaps its simplicity and low computational complexity. The LCA is essentially a delay-and-sum (DAS) beamformer where the delay step is carried out as usual, but where the weighting and summing is repeated for each of the $P$ windows. In its basic form its computational complexity is then of O($M\,P$).

@@

Some key aspects of our time domain MVDR implementation is illustrated in Figure \ref{mvdr_implementation}.

First we make each SM compute entire range lines of pixels, then load the subset of $\x$ that all these pixels depend upon into shared memory. This lets us perform temporal averaging without the need for reading additional channel data.

Second we assign $L$ threads per pixel to traverse the diagonals of $\eR$. On the top row a full computation is carried out, then that entire row is saved back to global memory following a collective access pattern that maximizes global memory bandwidth. For subsequent rows the threads move along the diagonals while performing iterative summations; the result from the previous element on the diagonal is updated by adding and removing the correlation coefficients that enters and exits the sum, respectively. To minimize memory consumption, we compute the coefficients again every time we need them. When a thread has finished up a diagonal, we have them wrap around to compute one of the diagonals in the lower triangular of $\eR$. Since $\eR$ is conjugate symmetric, the values in the leftmost column is obtained by a complex conjugate copy of the relevant value in the first row. Combined these steps balances the load evenly on all threads, is almost completely free of arithmetic redundancy, and consumes less memory.

@@

\begin{figure}[!h]\centering%
\graphicsAI[drawing,width=\linewidth]{gfx/plot_holmengraa_L16_Navg1.pdf}%
\caption{Holmengraa.}\label{mvdr_implementation}
\end{figure}

\begin{figure}[!h]\centering%
\graphicsAI[drawing,width=\linewidth]{gfx/benchmark_tagged.svg}%
\caption{Holmengraa.}\label{mvdr_implementation}
\end{figure}

>>>>

Although LCA promotes good image quality, its two most attractive traits is perhaps its simplicity and low computational complexity. The LCA is essentially a delay-and-sum (DAS) beamformer where the delay step is carried out as usual, but where the weighting and summing is repeated for each of the $P$ windows. In its basic form its computational complexity only of O($M\,P$).

@@

Our GPU implementation consists of three main kernels relating to the computing steps illustrated in Figure \ref{mvdr_implementation}. In the first step we estimate $\eR$, in the second we solve the linear equation $\eRi\1$, and finally we perform some substitutions to obtain the beamformer output, $z$. All pixels are processed before moving on to the next step. Our efforts here were mainly focused on estimating $\eR$, because an official batch linear equation solver is already available from Nvidia, and computing the beamformer output is a rather simple task.

When computing $\eR$ we we assign $L$ threads per pixel to traverse the diagonals of $\eR$. On the top row a full computation is carried out, then that entire row is saved back to global memory following a collective access pattern that maximizes global memory bandwidth. For subsequent rows the threads move along the diagonals while performing iterative summations; the result from the previous element on the diagonal is updated by adding and removing the correlation coefficients that enters and exits the sum, respectively. To minimize memory consumption, we compute the coefficients again every time we need them. When a thread has finished up a diagonal, we have them wrap around to compute one of the diagonals in the lower triangular of $\eR$. Since $\eR$ is conjugate symmetric, the values in the leftmost column is obtained by a complex conjugate copy of the relevant value in the first row. Combined these steps balances the load evenly on all threads, is almost completely free of arithmetic redundancy, and consumes less memory.

@@

We have tested our GPU implementation of the MVDR beamformer experimental datasets from the 32 element Kongsberg Maritime HISAS1030 sonar being attached to the HUGIN AUV~\cite{Hansen2009}. To obtain the studied object, the 1500 dwt oil tanker wreck Holmengraa, the sonar was operated in sidescan mode to image (\Fig{hlmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The image was processed with MVDR parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

The computational performance of our GPU implementation is displayed in (\Fig{benchmarks}). Processing times are shown for varying subarray lengths $L$ and varying amounts of temporal averaging $N_\text{avg}$. A single-thread C and Matlab implementation is added as reference, although only moderate optimization work was carried out on these. The test system was a quad-core Intel Xeon E5620, with 64Gb of RAM, and an nVidia Quadro 6000 card.

\begin{figure}[!t]\centering%
\graphicsAI[drawing,width=\linewidth]{gfx/plot_holmengraa_L16_Navg1.pdf}%
\caption{Holmengraa.}\label{holmengraa}
\end{figure}

\begin{figure}[!h]\centering%
\graphicsAI[drawing,width=\linewidth]{gfx/benchmark_tagged.svg}%
\caption{Holmengraa.}\label{benchmark}
\end{figure}
